\LARGE \bf Mean Estimation from Single-Bit Measurements

### 2018-09-06 Writeups/onebit_full_V2.tex
small change

%\documentclass[letterpaper]{IEEEtran}
\documentclass[letterpaper, conference, 11pt]{IEEEtran}      % Use this line for a4 paper

@@

%\onecolumn

\usepackage{mathptmx} 
\usepackage{times}

\usepackage{amsmath} 
\usepackage{amsbsy} 
\usepackage{amssymb}
%\usepackage{newtxtext, newtxmath}
\usepackage{mathrsfs}
\usepackage{comment}
\usepackage[export]{adjustbox}
\usepackage{tikz}
\usetikzlibrary{external,positioning,decorations.pathreplacing,shapes,arrows,patterns}

%\tikzexternalize[mode=list and make]
\usepackage{algorithmicx}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage{pstool}
\usepackage[latin1]{inputenc}
\usetikzlibrary{arrows,shapes}
\usepackage{xifthen}
\usepackage{epic}
\usepackage{caption}
\usepackage{epstopdf}

\newtheorem{thm}{\bf{Theorem}}
\newtheorem{cor}[thm]{\bf {Corollary}}
\newtheorem{lem}[thm]{\bf {Lemma}}
\newtheorem{prop}[thm]{\bf {Proposition}}
\newtheorem{example}{\bf {Example}}
\newtheorem{definition}{\bf {Definition}}
\newtheorem{rem}{\bf {Remark}}

\newcommand{\mmse}{\mathsf{mmse}}
\newcommand{\unif}{\mathsf{unif}}
\newcommand{\card}{\mathrm{card}}
\newcommand{\ARE}{\mathsf{ARE}}
\newcommand{\supp}{\mathrm{supp} }
\renewcommand\vec[1]{\ensuremath\boldsymbol{#1}}
\newenvironment{proof}{\paragraph*{Proof}}{\hfill$\square$ \newline}
\newcommand{\sgn}{\mathrm{sgn} }
\newcommand{\argmax}{\mathrm{argmax}}
\newcommand*{\QEDA}{\hfill\ensuremath{\square}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
We consider the problem of estimating the mean of a symmetric log-concave distribution under the constraint that only a single bit per sample from this distribution is available to the estimator. We study the mean squared error risk in this estimation as a function of the number of samples, and hence number of bits, from this distribution. Under an adaptive scheme in which each bit is a function of the current sample and the previously observed bits, we show that the optimal relative efficiency, compared to the standard mean estimator without bit limitation, is the efficiency of the median. For the distributed scheme we consider the setting where each bit is obtained by comparing against a prescribed threshold value. We show that the maximum likelihood estimator in this case is asymptotically local minimax, and its asymptotic relative efficiency is finite although strictly larger than that of the sample median. 
% and depends on the size of the parameter space from which the mean is taken. Furthermore, assuming a prior on this space, the efficiency in the distributed case approaches $\sigma^2/4f^2(0)$ provided the variance of the distribution is high compared to the variance of the prior. 
%
%Therefore, our results indicates that, rather surprisingly, parametric estimation from the coarsest form of quantization incurs only a multiplicative penalty factor on the number of measurements. For example, this factor is $\pi/2$ for the normal distribution. 
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:Intro}
Estimating parameters from data collected and processed by multiple units may be limited due to communication constraints between these units. 
%
For example, this scenario arises in sensor arrays where information is collected at multiple physical locations and transmitted to a central estimation unit. In these situations, the ability to estimate a particular parameter from the data is dictated not only by the quality of observations and their number, but also by the available bandwidth for communicating between the sensors and the central estimator. The question that we ask is to what extent a parametric estimation task is affected by this constraint on communication, and what are the fundamental performance limits in estimating a parameter subject to such restriction. \\

This paper answers this question in a particular setting: the estimation of the mean $\theta$ of a symmetric log-concave distribution $P_X$ with a finite variance, under the constraint that only a single bit can be communicated on each sample from this distribution. As it turns out, the ability to share information before committing on each single-bit message dramatically affects the performance in estimating $\theta$. We therefore distinguish among three settings:
\begin{itemize}
 \item[(i)] \emph{Centralized} encoding: all $n$ encoders confer and produce a single $n$ bit message. 
 \item[(ii)] \emph{Adaptive} or \emph{sequential} encoding: the $i$th encoder observes the $i$th sample and the $i-1$ previous messages.
 \item[(iii)] \emph{Distributed} encoding: the $i$th message is only a function of the $i$th observation.
 \end{itemize}
Evidently, as far as information sharing is concerned, settings (iii) is a more restrictive version of (ii) which is more restrictive than (i). Below are three application example of each of settings (i)-(iii), respectively:
\begin{itemize}
\item[(i)] {\bf Signal acquisition:} a quantity is measured $n$ times at different instances, and the results are averaged in order to reduce measurement noise. The averaged result is then stored using one of $n$ states. 
\item[(ii)] {\bf analog-to-digital conversion via sigma-delta modulation (SDM):} in SDM an analog signal is converted into a sequence of bits by sampling it at a very high rate and then using one-bit threshold detector combined with a feedback loop to update an accumulated error state. Therefore, the MSE in tracking an analog signal using a SDM falls under our setting (ii) when we assume that the signal at the input to the modulator is a constant (direct current) corrupted by, say, thermal noise \cite{53738}. Since the sampling rates in SDM is usually many times more than the bandwidth of its input, analyzing SDM under a constant input provides meaningful lower bound even for non-constant signals.
\item[(iii)] {\bf Differential privacy:} -- a business entity is interested in estimating the average income of its clients. In order to keep this information as confidential as possible, each client answers a single yes/no questions about its individual income. 
\end{itemize}

We measure the performance in estimating $\theta$ by the mean squared error (MSE) risk. We are interested in particular in the \emph{asymptotic relative efficiency} (ARE) of estimators in the constrained setting compared to asymptotically normal estimators whose variances decreases as $\sigma^2/n+o(n^{-1})$ where $\sigma^2$. Estimator of this form includes the empirical mean of $X_1,\ldots,X_n$, and under some conditions, optimal Bayes estimations. We note that the performance under estimating a signal that may vary with each noisy measurement is always harder that our setting where the parameter (the mean) remains fixed. Therefore, the excess MSE risk due to the one-bit per measurement constraint we derive in this paper can serve as the most optimistic estimate for the degradation in performance in estimating a signal that vary in time.

@@

In setting (i), the estimator can evaluate the empirical mean of the samples and then quantize it using $n$ bits. Since the accuracy in describing the empirical mean decreases exponentially in $n$, the error due to quantization is negligible compared to the MSE in estimating the mean. Therefore, the ARE in this setting is $1$. Namely, asymptotically, there is no loss in performance due to the communication constraint under centralized encoding. 
%
In this paper we show that a similar results does not hold even in setting and (ii): the ARE of any adaptive estimation scheme is at least that of the sample median. Specifically, when $P_X$ is normal, this ARE equals $2/\pi \approx 0.637$, showing that the one-bit constraint increases the effective sample size in estimating $\theta$ by at least $2/\pi \approx 1.57$ compared to estimating it without the bit constraint. We also show that this lower bound on the ARE and is tight by providing an estimator that attains it. % This estimator is obtained by averaging the results of a series of comparisons of the samples against certain thresholds. 
%This estimator is can be seen as a stochastic gradient descent procedure for minimizing the absolute error, hence it attains the asymptotic MSE of the sample median. % Interestingly, our estimator attains \eqref{eq:median_MSE} regardless of the particular realization of $\theta$ or the radius of the parameter space from which it is taken.
%
Clearly, the minimal penalty on the MSE in setting (ii) also holds under setting (iii), although the question whether this MSE is achievable (or otherwise, what is the minimal MSE) remains open. Instead, in setting (iii) we restrict ourselves to estimators from messages obtained by comparison against a prescribed value (that may be different for each sample). We show that the maximum likelihood (ML) estimator for $\theta$ from these messages is asymptotically local minimax, and its asymptotic variance is strictly greater than the variance of the median. Thus, at least when limited to threshold detection, the ability to adapt the threshold allows for a more efficient estimation. %This is in contrast to the \\ counterpart of our setting where $\theta$ is taken from a finite space

Even though the ARE in setting (i) is $1$, this scheme already poses a non-trivial challenge for the design and analysis of an optimal encoding and estimation procedures. Indeed, the standard technique to encode an unknown random quantity using $n$ bits is equivalent to the design of a scalar quantizer \cite{gray1998quantization}. However, the optimal design of this quantizer depends on the distribution of its input, which is the goal of our estimation problem and hence its exact value is unknown. As a result, a non-trivial exploration exploitation trade-off arises in this case. Note that the only missing parameter in our setting is the mean, which, under setting (i), is known to the encoder with uncertainty interval proportional to $\sigma/\sqrt{n}$. Therefore, while it is clear that uncertainty due to quantization decreases exponentially in the number of bits $n$ leading to ARE $1$, an exact expression for the MSE in this setting is still difficult to derive. \par
The situation is even more involved in the adaptive encoding setting (ii): an encoding and estimation strategy that is optimal for $n-1$ adaptive one-bit messages of a sample of size $n-1$, may not lead to a globally optimal strategy upon the recipient of the $n$th sample. Conversely, any one-step optimal strategy, in the sense that it finds the best one-bit message as a function of the current sample and the previous $n-1$ messages, is not guaranteed to be globally optimal. Therefore, while we characterize the optimal detector given the previous messages, this characterization cannot be use to derive a lower bound on the ARE. Instead, our result on the median efficiency being the minimal ARE is obtained by bounding the Fisher information of any $n$ adaptive messages and using the van Trees version of the information inequality \cite{gill1995applications}.

@@

\subsection*{Related Works}
As the variance $\sigma^2$ goes to zero, the task of finding $\theta$ using one-bit queries in the adaptive setting (ii) is easily solved by a bisection style method over the parameter space. Therefore, the general case of non-zero variance is a reminiscent of the noisy binary search problem with possibly infinite number of unreliable tests \cite{cicalese2002least, Karp:2007:NBS:1283383.1283478}. However, since we assume a continuous parameter space, a more closely related problem is that of one-bit analog-to-digital conversion of a constant input $\theta$ corrupted by a Gaussian noise. Using a SDM, Wong and Gray \cite{53738} showed that the output of the modulator converges to the true constant input almost surely, so that a SDM provides a consistent estimator for setting (ii). The rate of this convergence, however, was not analyzed and cannot be derived from the results of \cite{53738}. In particular, our results for setting (ii) imply that the asymptotic rate of convergence of the MSE in SDM to a constant input under Gaussian nosie is at most $\sigma^2\pi/2$ over the number of feedback iterations. Baraniuk et. al \cite{baraniuk2017exponential} also considered adaptive one-bit measurements in the context of analog-to-digital conversion, although without noise at the input. Their result of an exponential MSE decaying rate clearly does not hold in the noisy setting (stated otherwise, the MSE can decay exponentially up to the noise level). 
\\

One-bit measurements in the distributed setting (iii) was considered in \cite{904560,4244748, 6882252, chen2010performance, 5184907}, but without optimizing the encoders or their detection rule. % analyzed the estimation from distributed quantized measurements when the same detection rule is applied by all encoders. However, the error criterion considered in these works is the Fisher information rather than the MSE risk or ARE criterion. \par
%Once the decision rule of each encoder is fixed, the optimal estimation of $\theta$ is determined using the maximum a posteriori probability rule. Hence, the distributed setting is reduced to finding the optimal decision rule of each encoder.
The work of \cite{52470} addresses the counterpart of our setting (iii) in the case of hypothesis testing, although the results there cannot be extended to parametric estimation. More generally, when the parameter space $\Theta$ is finite, the characterization of the optimal detection rules was considered by Tsitsiklist in \cite{tsitsiklis1988decentralized}. It was shown there that if the cardinality of $\Theta$ is at most $M$ and the probability of error criterion is used, than no more than $M(M-1)/2$ different detection rules are necessary in order to attain probability of error decreasing exponentially with the optimal exponent. %That is, beyond $M(M-1)/2$ some decision rules can be repeated without lossing optimality.
Furthermore, in a version of this problem for the adaptive sestting \cite{5751320}, it was shown that, with a specific two-stages feedback, there is no gain in feedback compared to the fully distributed setting. Our results implies that the ARE in the distributed setting with threshold detection rules is strictly larger than that in the adaptive setting, suggesting that the case of a finite $\Theta$ is very different than when $\Theta$ is an open set.\par
%
As we explain in details in Section~\ref{sec:preliminary}, the remote multiterminal source coding problem, also known as the CEO problem \cite{berger1996ceo, viswanathan1997quadratic, oohama1998rate, prabhakaran2004rate}, leads to a lower bounds on the MSE in setting (iii). For the case of a Gaussian distribution, this lower bound bounds the ARE to be at most $3/4$. Thus, while this bound on the ARE provides no new information compared to the upper bound of $2/\pi$ we derive for setting (ii), it shows that the distributed nature of the problem is not a limiting factor in achieving MSE close to optimal even under one-bit quantization of each sample.\\

Finally, we note that our settings (ii) and (iii) can be obtained as special cases of 
\cite{zhang2013information} that consider adaptive and distributed estimation protocols for $m$ machines, each has access to $n/m$ independent samples. The main result of \cite{zhang2013information} are bounds on the estimation error as a function of the number of bits $R$ each machine uses for communication. The specialization of their result to our setting, by taking $m=n$ and $R=1$, leads to looser lower bounds then we derive here for cases (ii) and (iii). Other related works include statistical inference under multiterminal data compression \cite{han1987hypothesis, zhang1988estimation}, and one-bit quantization constraints 
%in compressed sensing\cite{baraniuk2017exponential} and 
in MIMO detection in wireless communication \cite{singh2009limits}. \\

@@

\node[int1, right of = enc2, node distance = 4cm ] (est) {Est};
\draw[->,line width = 2pt] (enc1) -| node[above, xshift = -2cm] (mes1) {$M_1 \in \left\{-1,1\right\}$} (est);   
\draw[->,line width = 2pt] (enc2) -- node[above, xshift = 0cm] (mes2) {$M_2 \in \left\{-1,1\right\}$} (est);   
\draw[->] (enc1)+(0.7,0) -- +(0.7,-0.7) -| (enc2);

@@

\node[right of = est, node distance = 1.5cm] (dest) {$\widehat{\theta}_n$};
%         \node [int1] (dec) [right of=dest, node distance = 1.5cm,  align=center] {\small Dec };
%\node [int1] (enc) [right of = dec, node distance = 3cm]{Enc}; 
%\draw[->,line width=2pt] (dec) -- (dest);
\draw[->, line width=2pt] (est) -- (dest);
\end{tikzpicture}
\end{center}
\caption{\label{fig:sequential} Adaptive single-bit encoding: the $i$th encoder delivers a single bit message which is a function of its private sample $X_i$ and the previous messages $M_1,\ldots,M_{i-1}$.}
\end{figure}

@@

\node[right of = est, node distance = 1.5cm] (dest) {$\widehat{\theta}_n$};
%         \node [int1] (dec) [right of=dest, node distance = 1.5cm,  align=center] {\small Dec };
%\node [int1] (enc) [right of = dec, node distance = 3cm]{Enc}; 
%\draw[->,line width=2pt] (dec) -- (dest);
\draw[->, line width=2pt] (est) -- (dest);
\end{tikzpicture}
\end{center}
\caption{\label{fig:distributed} Distributed single-bit encoding: the single-bit message produced by each encoder is only a function of its private sample $X_i$.}
\end{figure}
Let $\Theta$ be a closed interval of the real line. Let $f(x)$ be a symmetric log-concave density function with a finite second moment $\sigma^2$. We denote by $P_X$ the probability distribution with density $f\left( x-\theta \right)$, where $\theta \in \Theta$. Therefore, $P_X$ is an absolutely continuous log-concave distribution with mean $\theta$, variance $\sigma^2$. Moreover, symmetry and log-concavity of $f(x)$ implies that $P_X$ has a single mode at $x =\theta$. \par
In some cases we assume that $\theta$ is drawn once from the prior distribution $\pi$ on the parameter space $\Theta$. In this cases we assume that $\pi$  
is absolutely continuous distribution, and denote its density by $\pi(\theta)$, i.e., $\pi(d\theta) = \pi(\theta)d\theta$.  
\par
The random variables $X_1,\ldots,X_n$ represent $n$ independent samples from $P_X$. 
We are interested in estimating the parameter $\theta$ from a set of $n$ binary messages $M^n = (M_1,\ldots,M_n)$, obtained from $X^n = (X_1,\ldots,X_n)$ under three possible scenarios: 
\begin{itemize}
\item[(i)~~] Centralized $M^n(X^n)$ (Fig.~\ref{fig:centralized}).
\item[(ii)~] Adaptive $M_i(X_i,M^{i-1})$, $i=1,\ldots,n$ (Fig.~\ref{fig:sequential}).
\item[(iii)] Distributed $M_i(X_i)$, $i=1,\ldots,n$ (Fig.~\ref{fig:distributed}).
\end{itemize}

>>>>

%\documentclass[letterpaper]{IEEEtran}
\documentclass[letterpaper, 11pt]{IEEEtran}      % Use this line for a4 paper

@@

\onecolumn

\usepackage{mathptmx} 
\usepackage{times} 
\usepackage{amsmath} 
\usepackage{amsbsy} 
\usepackage{amssymb}
%\usepackage{newtxtext, newtxmath}
\usepackage{mathrsfs}
\usepackage{comment}
\usepackage[export]{adjustbox}
\usepackage{tikz}
\usetikzlibrary{external,positioning,decorations.pathreplacing,shapes,arrows,patterns}

%\tikzexternalize[mode=list and make]
\usepackage{algorithmicx}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage{pstool}
\usepackage[latin1]{inputenc}
\usetikzlibrary{arrows,shapes}
\usepackage{xifthen}
\usepackage{epic}
%\usepackage{caption}
\usepackage{epstopdf}

\newtheorem{thm}{\bf{Theorem}}
\newtheorem{cor}[thm]{\bf {Corollary}}
\newtheorem{lem}[thm]{\bf {Lemma}}
\newtheorem{prop}[thm]{\bf {Proposition}}
\newtheorem{example}{\bf {Example}}
\newtheorem{definition}{\bf {Definition}}
\newtheorem{rem}{\bf {Remark}}
\newtheorem{assump}{\bf {Assumptions}}

\newcommand{\mmse}{\mathsf{mmse}}
\newcommand{\unif}{\mathsf{unif}}
\newcommand{\card}{\mathrm{card}}
\newcommand{\ARE}{\mathsf{ARE}}
\newcommand{\supp}{\mathrm{supp} }
\renewcommand\vec[1]{\ensuremath\boldsymbol{#1}}
\newenvironment{proof}{\paragraph*{Proof}}{\hfill$\square$ \newline}
\newcommand{\sgn}{\mathrm{sgn} }
\newcommand{\Prob}{\mathbb{P} }
\newcommand{\argmax}{\mathrm{argmax}}
\newcommand*{\QEDA}{\hfill\ensuremath{\square}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
We consider the problem of estimating the mean of a symmetric log-concave distribution under the following constraint: only a single bit per sample from this distribution is available to the estimator. We study the mean squared error (MSE) risk in this estimation as a function of the number of samples, and hence number of bits, from this distribution. Under an adaptive setting in which each bit is a function of the current sample and the previously observed bits, we show that the optimal relative efficiency compared to the sample mean is the efficiency of the median. For example, in estimating the mean of a normal distribution, a constraint of one bit per sample incurs a penalty of $\pi/2$ in sample size compared to the unconstrained case. 
%
We also consider a distributed setting where each one bit message is only a function of a single sample. We derive lower bounds on the MSE in this setting, and show that the optimal efficiency can only be attained at a finite number of points in the parameter space. 
%
Finally, we analyze a distributed setting where the bits are obtained by comparing each sample against a prescribed threshold, and consider the threshold density that minimizes the maximal MSE. 
%
Our results indicates that estimating the mean from one bit measurements is equivalent to estimating the sample median from these measurements. In the adaptive case, this estimate can be done with vanishing error for any point in the parameter space. In the distributed case, this estimate can be done with vanishing error only for a finite number of possible values for the unknown mean. 
%
%Therefore, our results indicates that, rather surprisingly, parametric estimation from the coarsest form of quantization incurs only a multiplicative penalty factor on the number of measurements. For example, this factor is $\pi/2$ for the normal distribution. 
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:Intro}
Estimating parameters from data collected and processed by multiple units may be limited due to communication constraints between these units. 
%
For example, this scenario arises in sensor arrays where information is collected at multiple physical locations and transmitted to a central estimation unit. In these situations, the ability to estimate a particular parameter from the data is dictated not only by the quality of observations and their number, but also by the available bandwidth for communicating between the sensors and the central estimator. The question that we ask is to what extent a parametric estimation task is affected by this constraint on communication, and what are the fundamental performance limits in estimating a parameter subject to such restriction. \par
This paper answers this question in a particular setting: the estimation of the mean $\theta$ of a symmetric log-concave distribution with a finite variance, under the constraint that only a single bit can be communicated on each sample from this distribution. As it turns out, the ability to share information before committing on each single-bit message dramatically affects the performance in estimating $\theta$. We therefore distinguish among three settings:
\begin{itemize}
 \item[(i)\,\,] \emph{Centralized} encoding: all $n$ encoders confer and produce a single $n$ bit message. 
 \item[(ii)\,] \emph{Adaptive} or \emph{sequential} encoding: the $n$th encoder observes the $n$th sample and the $n-1$ previous messages (bits).
 \item[(iii)] \emph{Distributed} encoding: the $n$th message is only a function of the $n$th observation.
 \end{itemize}
Evidently, as far as information sharing is concerned, settings (iii) is a more restrictive version of (ii) which is more restrictive than (i). Below are three application examples for each of settings (i)-(iii) above, respectively:
\begin{itemize}
\item {\bf Signal acquisition:} a quantity is measured $n$ times at different instances, and the results are averaged in order to reduce measurement noise. The averaged result is then stored using one of $n$ states. 
\item {\bf Analog-to-digital conversion:} in sigma-delta modulation, an analog signal is converted into a sequence of bits by sampling it at a very high rate and then using one-bit threshold detector combined with a feedback loop to update an accumulated error state. Therefore, the MSE in tracking an analog signal using a SDM falls under our setting (ii) when we assume that the signal at the input to the modulator is a constant (direct current) corrupted by, say, thermal noise \cite{53738}. Since the sampling rates in SDM are usually many times more than the bandwidth of its input, analyzing SDM under a constant input provides meaningful lower bound even for non-constant signals.
\item {\bf Differential privacy:} -- a business entity is interested in estimating the average income of its clients. In order to keep this information as confidential as possible, each client independently provides an answer to a yes/no question related to its income.
\end{itemize}
%
We measure the performance in estimating $\theta$ by the mean squared error (MSE) risk. We are interested in particular in the \emph{asymptotic relative efficiency} (ARE) of estimators in the constrained setting compared to asymptotically normal estimators whose variances decreases as $\sigma^2/n+o(n^{-1})$. Estimators of this form include the empirical mean of the samples, and, under some conditions, the optimal Bayes estimator. \par
%
In addition to the examples above, the excess risk in estimating a fixed parameter due to a one-bit per measurement constraint is useful in bounding from below the excess risk in estimating a signal from its noisy measurements. Namely, the excess MSE or ARE we derive serves as the most optimistic estimate for the risk in estimating a signal that varies in time or space under the one bit per measurement constraint. Such situtations are considered in \cite{baraniuk2017exponential, jacques2013robust, plan2013one, li2017channel, choi2016near}. \\

@@

In setting (i), the estimator can evaluate the optimal mean estimator (e.g., the sample mean in the Gaussian case) and then quantize it using $n$ bits. Since the accuracy in describing the empirical mean decreases exponentially in $n$, the error due to quantization is negligible compared to the MSE in estimating the mean. Therefore, the ARE in this setting is $1$. Namely, asymptotically, there is no loss in performance due to the communication constraint under centralized encoding. 
%
In this paper we show that a similar results does not hold in setting (ii): the ARE of any adaptive estimation scheme is at most the ARE of the sample median. Specifically, when the samples are drawn from the normal distribution, this ARE equals $\pi/2$, showing that the one-bit constraint increases the effective sample size in estimating $\theta$ by at least $\pi/2 \approx 1.57$ compared to estimating it without the bit constraint. We also show that this lower bound on the ARE is tight by providing an estimator that attains it. % This estimator is obtained by averaging the results of a series of comparisons of the samples against certain thresholds. 
%This estimator is can be seen as a stochastic gradient descent procedure for minimizing the absolute error, hence it attains the asymptotic MSE of the sample median. % Interestingly, our estimator attains \eqref{eq:median_MSE} regardless of the particular realization of $\theta$ or the radius of the parameter space from which it is taken.
%
Clearly, the penalty on the sample size in setting (iii) is at least as large as that in setting (ii). 
Unlike in setting (ii), we show that there is no distributed estimation scheme that is uniformly optimal in the sense that it attains the ARE of the sample median for all $\theta$ in the parameter space. 
% Instead, in setting (iii) we restrict ourselves to estimators from messages obtained by comparison against a prescribed value (that may be different for each sample). We show that the maximum likelihood (ML) estimator for $\theta$ from these messages is asymptotically local minimax, and its asymptotic variance is strictly greater than the variance of the median. Thus, at least when limited to threshold detection, the ability to adapt the threshold allows for a more efficient estimation. %This is in contrast to the \\ counterpart of our setting where $\theta$ is taken from a finite space

We note that although the ARE in setting (i) is $1$, this scheme already poses a non-trivial challenge for the design and analysis of an optimal encoding and estimation procedures. Indeed, the standard technique to encode an unknown random quantity using $n$ bits is equivalent to the design of a scalar quantizer \cite{gray1998quantization}. However, the optimal design of this quantizer depends on the distribution of its input, which is the goal of our estimation problem and hence its exact value is unknown. As a result, a non-trivial exploration exploitation trade-off arises in this case. 
%Note that the only missing parameter in our setting is the mean, which, under setting (i), is known to the encoder with uncertainty interval proportional to $\sigma/\sqrt{n}$. 
Therefore, while uncertainty due to quantization decreases exponentially in the number of bits $n$, hence the ARE is $1$, an exact expression for the MSE in this setting is still difficult to derive. \par
%
The situation is even more involved in the adaptive encoding setting (ii): an encoding and estimation strategy that is optimal for $n-1$ adaptive one-bit messages of a sample of size $n-1$ may not lead to a globally optimal strategy upon the recipient of the $n$th sample. Conversely, any one-step optimal strategy, in the sense that it finds the best one-bit message as a function of the current sample and the previous $n-1$ messages, is not guaranteed to be globally optimal. Therefore, while we characterize the optimal one-bit message given the previous messages, this characterization does not necessarily  leads to an upper bound on the ARE. Instead, our result on the maximal ARE is obtained by bounding the Fisher information of any $n$ adaptive messages and using an appropriate information inequality. \par
In addition to encoding and estimation schemes that leads to optimal results, we also consider two additional ``natural'' schemes. Specifically, in setting (ii) we consider the one-bit optimal scheme, i.e., the case of a greedy encoder that given the $n$th sample and the previous $n-1$ bits, provides a bit that minimizes the $n$-step MSE. In setting (iii) we also consider the case where the messages are obtained by comparing each sample against a prescribed threshold. This threshold may be different across samples and is assumed deterministic (independent of the data).

@@

\subsection*{Related Works}
As the variance $\sigma^2$ goes to zero, the task of finding $\theta$ using one-bit queries in the adaptive setting (ii) is solved by a bisection style method over the parameter space. Therefore, the general case of non-zero variance is a reminiscent of the noisy binary search problem with possibly infinite number of unreliable tests \cite{cicalese2002least, Karp:2007:NBS:1283383.1283478}. However, since we assume a continuous parameter space, a more closely related problem is that of one-bit analog-to-digital conversion of a constant input corrupted by a Gaussian noise. Using a SDM, Wong and Gray \cite{53738} showed that the output of the modulator converges to the true constant input almost surely, so that a SDM provides a consistent estimator for setting (ii). The rate of this convergence, however, was not analyzed and cannot be derived from the results of \cite{53738}. In particular, our results for setting (ii) imply that the asymptotic rate of convergence of the MSE in SDM to a constant input under an additive white Gaussian nosie is at most $\sigma^2\pi/2$ over the number of feedback iterations. Baraniuk et. al \cite{baraniuk2017exponential} also considered adaptive one-bit measurements in the context of analog-to-digital conversion, although without noise at the input. By establishing the lower bound of $\sigma^2\pi/2n$ on the MSE, we show that the main results of \cite{baraniuk2017exponential}, an exponential MSE decaying rate, does not hold in the noisy setting. Stated otherwise, the MSE in the setting of \cite{baraniuk2017exponential} may decay exponentially up to the noise level, after which it decays at most as $\sigma^2\pi/2n$. \par
%
One-bit measurements in the distributed setting (iii) was considered in \cite{904560,4244748, 6882252, chen2010performance, 5184907}, but without optimizing the encoders and their detection rules. Consequently, the performance derived in these works are not optimal. 
 % analyzed the estimation from distributed quantized measurements when the same detection rule is applied by all encoders. However, the error criterion considered in these works is the Fisher information rather than the MSE risk or ARE criterion. \par
%Once the decision rule of each encoder is fixed, the optimal estimation of $\theta$ is determined using the maximum a posteriori probability rule. Hence, the distributed setting is reduced to finding the optimal decision rule of each encoder.
The work of \cite{52470} addresses the counterpart of our setting (iii) in the case of hypothesis testing, although the results there cannot be extended to parametric estimation. When the parameter space $\Theta$ is finite, Tsitsiklist \cite{tsitsiklis1988decentralized} showed that when the cardinality of $\Theta$ is at most $M$ and the probability of error criterion is used, then no more than $M(M-1)/2$ different detection rules are necessary in order to attain probability of error decreasing exponentially with the optimal exponent. %That is, beyond $M(M-1)/2$ some decision rules can be repeated without lossing optimality.
Furthermore, in a version of this problem for the adaptive sestting \cite{5751320}, it was shown that, with a specific two-stages feedback, there is no gain in feedback compared to the fully distributed setting. Our results implies that the ARE in the distributed setting with threshold detection rules is strictly larger than that in the adaptive setting, suggesting that the case of a finite $\Theta$ is very different than the case where $\Theta$ is an open set.\par
%
As we explain in detail in Section~ \ref{sec:preliminary}, the remote multiterminal source coding problem, also known as the CEO problem \cite{berger1996ceo, viswanathan1997quadratic, oohama1998rate, prabhakaran2004rate}, leads to a lower bounds on the MSE in setting (iii). For the case of a Gaussian distribution, this lower bound bounds the ARE to be at most $3/4$. Thus, while this bound on the ARE provides no new information compared to the upper bound of $2/\pi$ we derive for setting (ii), it shows that the distributed nature of the problem is not a limiting factor in achieving MSE close to optimal even under the one bit per sample constraint. 
 \par
% 
Finally, we note that our settings (ii) and (iii) can be obtained as special cases of 
\cite{zhang2013information} 
that consider adaptive and distributed estimation protocols for $m$ machines, each has access to $n/m$ independent samples. The main result of \cite{zhang2013information} are bounds on the estimation error as a function of the number of bits $R$ each machine uses for communication. The specialization of their result to our setting, by taking $m=n$ and $R=1$, leads to looser lower bounds then we derive here for cases (ii) and (iii). Looser bounds can also be obtained from \cite{zhang1988estimation, han2018distributed} that considered inference and distributed estimation under data compression constraint \cite{DBLP:journals/corr/abs-1802-08417, zhang1988estimation, han2018distributed}. \\
%in MIMO detection in wireless communication \cite{singh2009limits}. \\

The remainder of this paper is organized as follows. In Section~\ref{sec:problem} we describe the problem and useful notation. In Section~\ref{sec:preliminary} we provide two simple bounds on the efficiency and MSE. Our main results for the adaptive and distributed cases are given in Sections~\ref{sec:sequential} and \ref{sec:distributed}, respectively. In Section~\ref{sec:conclusions} we provide concluding remarks.

@@

\node[int1, right of = enc2, node distance = 4cm ] (est) {Est};
\draw[->,line width = 2pt] (enc1) -| node[above, xshift = -2cm] (mes1) {$M_1 \in \left\{-1,1\right\}$} (est);   
\draw[->,line width = 2pt] (enc2) -- node[above, xshift = 0cm] (mes2) {$M_2 \in \left\{-1,1\right\}$} (est);

\draw[->] (enc1)+(0.7,0) -- +(0.7,-0.7) -| (enc2);

@@

\node[right of = est, node distance = 1.5cm] (dest) {$\hat{\theta}_n$};
%         \node [int1] (dec) [right of=dest, node distance = 1.5cm,  align=center] {\small Dec };
%\node [int1] (enc) [right of = dec, node distance = 3cm]{Enc}; 
%\draw[->,line width=2pt] (dec) -- (dest);
\draw[->, line width=2pt] (est) -- (dest);
\end{tikzpicture}
\end{center}
\caption{\label{fig:sequential} Adaptive single-bit encoding: the $n$th encoder delivers a single bit message which is a function of its private sample $X_n$ and the previous messages $M_1,\ldots,M_{n-1}$.}
\end{figure}

@@

\node[right of = est, node distance = 1.5cm] (dest) {$\hat{\theta}_n$};
%         \node [int1] (dec) [right of=dest, node distance = 1.5cm,  align=center] {\small Dec };
%\node [int1] (enc) [right of = dec, node distance = 3cm]{Enc}; 
%\draw[->,line width=2pt] (dec) -- (dest);
\draw[->, line width=2pt] (est) -- (dest);
\end{tikzpicture}
\end{center}
\caption{\label{fig:distributed} Distributed single-bit encoding: the single-bit message produced by each encoder is only a function of its private sample $X_i$.}
\end{figure}
Let $f(x)$ be a symmetric and log-concave density function with a finite second moment $\sigma^2$. For $\theta \in \Theta$, denote by $P_X$ the probability distribution with density $f\left( x-\theta \right)$. Therefore, $P_X$ is an absolutely continuous log-concave distribution with mean $\theta$ and variance $\sigma^2$. Symmetry and log-concavity of $f(x)$ imply that $P_X$ is strongly unimodal with its mode at $x =\theta$ \cite{ibragimov1956composition}. We further assume that the \emph{parameter space} $\Theta$ is a closed interval of the real line. 
\par
In some situations it is useful to assume that $\theta$ is drawn once from the prior distribution $\pi$ on $\Theta$. In this case we assume that $\pi$ is an absolutely continuous distribution, and denote its density by $\pi(\theta)$, i.e., $\pi(d\theta) = \pi(\theta)d\theta$.  
\par
The random variables $X_1,\ldots,X_n$ represent $n$ independent samples from $P_X$. 
We are interested in estimating $\theta$ from a set of $n$ binary messages $M_1,\ldots,M_n$, obtained from $X_1,\ldots,X_n$ under three possible scenarios: 
\begin{itemize}
\item[(i)~~] Centralized $M_i(X_1,\ldots,X_n)$, $i=1,\ldots,n$ (Fig.~\ref{fig:centralized}).
\item[(ii)~] Adaptive $M_i(X_i,M_1,\ldots,M_{i-1})$, $i=2,\ldots,n$ (Fig.~\ref{fig:sequential}).
\item[(iii)] Distributed $M_i(X_i)$, $i=1,\ldots,n$ (Fig.~\ref{fig:distributed}).
\end{itemize}
